VIPE: Virtualized Intermittently-Powered Executor

### 2017-07-25 paper/background.tex
did some revision in the background

Performing complex operations on battery-less energy harvesting embedded platforms is challenging from many points of view. Here we shall briefly provide a background on computation with energy harvesting embedded systems.

\subsection{Energy Harvesting: Background}
\label{sec:background_harvesting}

Supplying power to tiny embedded computers using batteries alone is not sustainable. European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}. Since not all batteries are recycled, they constitute a potential environmental hazard. A potent solution is to replace batteries with smaller, but environment friendly {(super-)capacitors} (promising millions of charge/discharge cycles~\cite[Sec. I]{ongaro_pwre_2012}) and to supply them {\em directly from the ambient energy sources}---creating a truly self-sustainable computing device. The limitation of the capacitor as a storage device is energy density that is inferior to a battery, making effective capacitor discharge times orders of magnitude smaller than for a battery.

Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible. These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.

Many platforms enable intermittent, battery-less computation. For instance, computation RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}. Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}. In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}). Naturally, other forms of energy harvesting sources exist, including temperature gradient, (micro-)motions, light/sun radiation, vibrations, and body fluid flow (blood, gastric acid). The reader is referred to example extensive surveys discussing energy harvesting for low-power embedded systems~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016}.

@@

\subsection{Memory Consistency}
\label{sec:background_consistency}

>>>>

\sys addresses the problem of intermittently executing task-based computations
on energy-harvesting hardware, without being limited by inflexible, static task
decomposition and without being limited to fully non-volatile main memory.  In
this section, we provide background on energy harvesting systems and the
intermittent software execution model.   We then discuss the key limitations of
existing task-based programming and execution models for intermittent computing
that \sys addresses.

\subsection{Energy Harvesting Systems}
\label{sec:background_harvesting}

Energy harvesting devices are systems that operate using energy extracted from
their environment, from sources such as radio waves (RF)~\cite{wisp}, and solar
energy.  By operating entirely using environmental energy, energy-harvesting
devices need not use a tethered power supply or battery.  Instead, devices
typically collect energy into a (super-)capacitor, operate when sufficient
energy has accumulated, and upon depleting the stored energy, turn off and wait
to collect more energy.

Batteryless operation has a number of important advantages, making intermittent
computing an important domain.  Supplying power to the ``next
billion''~\cite{citationneeded} embedded computers using batteries is not
sustainable.  The European Commission estimates that more than 160 kilotons of
consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
Batteries constitute a potential environmental hazard because not all batteries
are recycled.  Batteries are fragile and limited in their number of
charge/discharge cycles, requiring costly maintenance that is possibly only
through physical device access, which may be difficult if devices are deployed
in inaccessible places (e.g., in space~\cite{kicksat}). By contrast,
super-capacitors are durable, promising millions of charge/discharge
cycles~\cite[Sec. I]{ongaro_pwre_2012} The main limitation in moving to
capacitor-based energy storage is that capacitor energy density --- and
consequently operating discharge time --- is orders of magnitude less than a
battery.

\TODO{I think this can be compressed a bit}
Given current technology development, battery-less systems are best suited for
very long-term sensing and monitoring where access to recharge is either
prohibitive or impossible. These include battery-less image capture and
processing~\cite{naderiparizi_rfid_2015}, animal
monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}
and digestible~\cite{nadeau_naturebio_2017} sensors.
% 
Many platforms enable intermittent, battery-less computation. For instance,
computation RFIDs---open-source TI MSP430-based~\cite{wolverine}
WISP~\cite{wisp5} (with its variants such as
WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or
NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones
such as~\cite{medusa_farsens_2017}. Other intermittently-powered platforms
include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or
battery-less phone~\cite{talla_imwut_2017}. In all of the above, the main
source of energy harvested is the electromagnetic radiation in the radio
frequency range (ambient transmitters such as high power TV
transmitters~\cite{liu_sigcomm_2013} or dedicated RFID
antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
Naturally, other forms of energy harvesting sources exist, including
temperature gradient, (micro-)motions, light/sun radiation, vibrations, and
body fluid flow (blood, gastric acid).  Several recent surveys discussing
energy harvesting, low-power, embedded systems and intermittent computing at a
high
level~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016,lucia_snapl_2017}.

@@

\subsection{Intermittent Computing}
\label{sec:background_consistency}
